---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# moodleStats

<!-- badges: start -->

[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental) [![R-CMD-check](https://github.com/Lightbridge-KS/moodleStats/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/Lightbridge-KS/moodleStats/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/Lightbridge-KS/moodleStats/branch/main/graph/badge.svg)](https://app.codecov.io/gh/Lightbridge-KS/moodleStats?branch=main)
<!-- badges: end -->

> A high-level R package for quiz & questions analysis of [Moodle Grades Report](https://docs.moodle.org/311/en/Quiz_reports#Grades_report)

# Installation

You can install the development version of moodleStats from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("Lightbridge-KS/moodleStats")
```

# Goal

The goal of this package it to provide a high-level functions for analysis of [Moodle Grades Report](https://docs.moodle.org/311/en/Quiz_reports#Grades_report) such as calculation of **descriptive statistics** for quiz & questions, and performingan **item analysis**.

# Workflow

```{r load-pkg}
library(moodleStats)
```

## Read Data

Read Moodle Grades Report from `.csv` file into a Data Frame.

```{r grades_df}
grades_df <- readr::read_csv(moodleStats_example("grades_report.csv"))
```

```{r}
head(grades_df)
```

## Prepare Data

Cleaning and filtering data can be done in 1 step using `prep_grades_report()`

```{r grades_df_preped}
grades_df_preped <- prep_grades_report(grades_df,
                                       choose_state = "Finished", # 1. choose only "Finished" attempt
                                       choose_grade = "max", # 2. choose only the best score of each student
                                       choose_time = "first" # 3. choose the first time that student submitted
                                       ) 
head(grades_df_preped)
```

## Quiz Metadata

Show quiz metadata such as maximum settings of quiz and questions by `quiz_meta()`

```{r grades_df_meta}
quiz_meta(grades_df_preped)
```

## Quiz Summary

`summary_quiz()` calculates various quiz summary statistics, and combine it into 1 row data frame.

```{r quiz_report_df}
quiz_report_df <- summary_quiz(grades_df_preped)
quiz_report_df
```

## Question Summary and Item Analysis

`summary_questions()` calculates various question summary statistics and item analysis into a data frame.

```{r question_report_df}
question_report_df <- summary_questions(grades_df_preped)
```

Here is summary statistics for each questions.

```{r}
question_report_df %>% 
  dplyr::select(Questions:SD)
```

And, here is an **item analysis** for Moodle Grades Report.

-   **`Difficulty_Index`** is for **Item Difficulty Index (p)**, a measure of the *proportion* of examinees who answered the item correctly.
-   **`Discrimination_Index`** is for **Item Discrimination Index (r)**, a pairwise *point-biserial correlation* between the score of each questions and total score of the quiz.

```{r}
question_report_df %>% 
  dplyr::select(Questions, Difficulty_Index:p.value)
```

# Learn more

-  [Get started using moodleStats](https://lightbridge-ks.github.io/moodleStats/articles/moodleStats.html)

------------------------------------------------------------------------

Last updated: `r format(Sys.time(), '%d %B %Y')`
